---
title: "Plate Discipline"
author: "Joshua Mould"
date: "2/28/2021"
output: html_document
---

# Quantifying MLB Hitter Plate Discipline

Import Necessary Packages
```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(baseballr)
library(gam)
library(randomForest)
library(caret)
library(xgboost)
library(mlr)
library(ggthemes)
library(e1071)
#install.packages("remotes")

options(scipen = 999)


GetAUC <- function(model, data, Target){
library(pacman)
p_load(ROCR)

pred <- predict(model, type="response")
  prediction <- prediction(pred, Target)
  Perf <- performance(prediction, "auc")
  return(as.numeric(Perf@y.values[1]))

}
```


Import MLB 2019 and 2020 pitch by pitch data
```{r}
hit_data <- readRDS("MLB Pitches 2019-2020.rds")
pitchers_2019 <- readRDS("Statcast Pitchers Pitch Data 2019.rds")
pitchers_2020 <- readRDS("Statcast Pitchers Pitch Data 2020.rds")
pitchers <- rbind(pitchers_2019, pitchers_2020)
pitchers <- pitchers %>% select(game_pk, at_bat_number, pitch_number, player_name) %>% rename(pitcher_name = player_name)
hit_data <- hit_data %>% left_join(pitchers, by = c("game_pk", "at_bat_number", "pitch_number"))



hit_data <- hit_data %>% select(-spin_dir, -spin_rate_deprecated, -break_angle_deprecated, -break_length_deprecated, -tfs_deprecated, -tfs_zulu_deprecated, -fielder_2, -umpire, -vx0, -vy0, -vz0, -ax, -ay, -az, -pitcher_1, -fielder_2_1, -fielder_3, -fielder_4, -fielder_5, -fielder_6, -fielder_7, -fielder_8, -fielder_9, -if_fielding_alignment, -of_fielding_alignment)
```

Add run expectancy metrics 
```{r}
hit_data <- hit_data %>% run_expectancy_code(level = "pitch")
hit_data$description <- as.factor(hit_data$description)
```

## Ball/Strike Model
Fit a glm model to determine balls and strikes
```{r}

takes <- hit_data %>% filter(description == "ball" | description == "called_strike", !is.na(plate_x) & !is.na(plate_z)) %>% mutate(strike = ifelse(description == "called_strike", 1, 0),
         plate_x2 = plate_x^2, plate_z2 = plate_z^2,
         plate_x3 = plate_x2 * plate_x, plate_z3 = plate_z2 * plate_z,
         plate_xz = plate_x * plate_z,
         plate_x2z = plate_x^2 * plate_z,
         plate_xz2 = plate_x * plate_z^2)

strike_prob_model <- glm(strike ~ plate_x + plate_z + plate_x2 + plate_z2 + plate_xz + sz_top + sz_bot, data = takes, family = binomial(link = "logit"))

summary(strike_prob_model)

#takes %>% add_predictions(strike_prob_model, type = "response")

GetAUC(strike_prob_model, takes, takes$strike)
#takes %>% select(strike, plate_x, plate_z, plate_x2,  plate_z2, plate_x3, plate_z3, plate_xz, plate_x2z, plate_xz2) %>% cor()


```

Create a strike heatmap to see the accuracy of the logistic model
```{r}
x <- seq(-1.5, 1.5, length.out=100)
y <- seq(0.5, 5, length.out=100)
preds <- data.frame(plate_x = c(outer(x, y * 0 + 1)), plate_z = c(outer(x * 0 + 1, y)), sz_top = mean(hit_data$sz_top, na.rm = T), sz_bot =mean(hit_data$sz_bot, na.rm = T))

preds <- preds %>% mutate(plate_x2 = plate_x^2, plate_z2 = plate_z^2,
         plate_x3 = plate_x2 * plate_x, plate_z3 = plate_z2 * plate_z,
         plate_xz = plate_x * plate_z,
         plate_x2z = plate_x^2 * plate_z,
         plate_xz2 = plate_x * plate_z^2) %>% add_predictions(strike_prob_model) %>% mutate(strike_prob = 1/(1+exp(-1*pred)))

#strike_predictions <- predict(strike_prob_model, type="response", newdata = preds)
#preds <- preds %>% add_column(strike_predictions)

topKzone <- mean(hit_data$sz_top, na.rm = T)
botKzone <- mean(hit_data$sz_bot, na.rm = T)
inKzone <- -0.84
outKzone <- 0.84
kZone <- data.frame(
  x=c(inKzone, inKzone, outKzone, outKzone, inKzone),
  y=c(botKzone, topKzone, topKzone, botKzone, botKzone))

# construct the plot

ggplot(kZone, aes(x, y)) +
    geom_tile(data=preds, 
              aes(x=plate_x, y=plate_z, fill= strike_prob)) +
    scale_fill_distiller(palette = "Spectral") +
    geom_path(lwd=1.5, col="black") +
    coord_fixed() +
    ggtitle("Ball vs Strike") +
  labs(fill = "Strike Probability", y = "z")


```


Logic for run expectancy ball and strike data frames, create run expectancy states for next possible outcomes (ball/strike) based on current situations.
```{r}
takes$pitch_id <- seq(1, nrow(takes))

ball_next <- takes
ball_next <- ball_next %>% mutate(balls_new = ifelse(balls < 3, balls + 1, 0), 
                                  strikes_new = ifelse(balls < 3, strikes, 0), 
                                  outs_when_up_new = outs_when_up, 
                                  on_1b_new = ifelse(balls == 3 & is.na(on_1b), 1, on_1b), 
                                  on_2b_new = ifelse(balls == 3 & !is.na(on_1b), 1, on_2b), 
                                  on_3b_new = ifelse(balls == 3 & !is.na(on_1b) & !is.na(on_2b), 1, on_3b)) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
ball_next <- ball_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))

strike_next <- takes
strike_next <- strike_next %>% mutate(balls_new = ifelse(strikes < 2, balls, 0), 
                                  strikes_new = ifelse(strikes < 2, strikes + 1, 0), 
                                  outs_when_up_new = ifelse(strikes < 2, outs_when_up, outs_when_up + 1), 
                                  on_1b_new =  on_1b, 
                                  on_2b_new = on_2b, 
                                  on_3b_new = on_3b) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
strike_next <- strike_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))

strike_next[which(strike_next$outs_when_up_new == 3), "avg_re.y"] <- 0

```

Combine the resulting data frames and calculate strike probability
```{r}
takes_ball <- ball_next %>% 
  select(pitch_id, balls_new, strikes_new, outs_when_up_new, on_1b_new, on_2b_new, on_3b_new, count_base_out_state_new, avg_re.y) %>% 
  rename(count_base_out_state_ball = count_base_out_state_new, avg_re.ball = avg_re.y)

take_outcomes <- strike_next %>% 
  rename(count_base_out_state_orig = count_base_out_state, count_base_out_state_strike = count_base_out_state_new, avg_re.orig = avg_re.x, avg_re.strike = avg_re.y) %>% left_join(takes_ball, by = "pitch_id")

take_outcomes <- take_outcomes %>% add_predictions(strike_prob_model, var = "strike_prob", type = "response")

```

Create variable with the run expectancy total for taking a pitch
```{r}
take_outcomes <- take_outcomes %>% mutate(re_take = avg_re.ball * (1 - strike_prob) - avg_re.strike * strike_prob)


take_outcomes %>% select(plate_x, plate_z, balls, strikes, on_1b, on_2b, on_3b, strike_prob, avg_re.strike, avg_re.ball, re_take) %>% arrange(re_take) %>% head()


```


## Build xgboost Multi-Classification Probability Model
Use statcast data for creating metrics for players with the swings dataset to be used in the classification model. 
```{r, echo = F, results=F, warning=FALSE}

basic_batter_stats <- daily_batter_bref("2019-03-20", "2020-09-27")

## merge to data 
hit_data <- hit_data %>% inner_join(basic_batter_stats, by = c("player_name" = "Name"))

hit_data <- hit_data %>% mutate(barrel = ifelse(launch_angle <= 50 & launch_speed >= 98 & launch_speed * 1.5 - launch_angle >= 117 & launch_speed + launch_angle >= 124, 1, 0))


# Fix character columns
hit_data[which(hit_data$launch_speed_angle == "null"),"launch_speed_angle"] <- NA
hit_data[which(hit_data$estimated_ba_using_speedangle == "null"), "estimated_ba_using_speedangle"] <- NA
hit_data[which(hit_data$estimated_woba_using_speedangle == "null"), "estimated_woba_using_speedangle"] <- NA
hit_data[which(hit_data$woba_value == "null"), "woba_value"] <- NA
hit_data[which(hit_data$iso_value == "null"), "iso_value"] <- NA

hit_data$launch_speed_angle <- as.numeric(hit_data$launch_speed_angle)
hit_data$estimated_ba_using_speedangle <- as.numeric(hit_data$estimated_ba_using_speedangle)
hit_data$estimated_woba_using_speedangle <- as.numeric(hit_data$estimated_woba_using_speedangle)
hit_data$woba_value <- as.numeric(hit_data$woba_value)
hit_data$iso_value <- as.numeric(hit_data$iso_value)


grouped_stats <- hit_data %>% group_by(player_name, zone) %>% summarise(barrel_perc = mean(barrel, na.rm = T), avg_launch_angle = mean(launch_angle, na.rm = T), avg_exit_velocity = mean(launch_speed, na.rm = T), avg_launch_speed_angle = mean(launch_speed_angle, na.rm = T), xAVG = mean(estimated_ba_using_speedangle, na.rm = T), xwOBA = mean(estimated_woba_using_speedangle, na.rm = T), wOBA = mean(woba_value, na.rm = T), ISO = mean(iso_value, na.rm = T))


hit_data <- hit_data %>% inner_join(grouped_stats, by = c("player_name", "zone"))

hit_data <- hit_data %>% separate(pitcher_name, into = c("last_name", "first_name"), sep = ", ") %>% mutate(pitcher_name = paste(first_name, last_name, sep = " "))

basic_pitcher_stats <- daily_pitcher_bref("2019-03-20", "2020-09-27")

basic_pitcher_stats$bbref_id <- as.numeric(basic_pitcher_stats$bbref_id)

hit_data <- hit_data %>% inner_join(basic_pitcher_stats, by = c("pitcher_name" = "Name"))

grouped_pitching_stats <- hit_data %>% group_by(pitcher_name) %>% summarise(FF_spin_rate = mean(release_spin_rate[pitch_type == "FF"], na.rm = T), CU_spin_rate = mean(release_spin_rate[pitch_type == "CU"], na.rm = T), FF_velocity = mean(release_speed[pitch_type == "FF"], na.rm = T), BRK_velocity = mean(release_speed[pitch_type == "CU" | pitch_type == "SL"], na.rm = T), overall_spin_rate = mean(release_spin_rate, na.rm = T), overall_velocity = mean(release_speed, na.rm = T))

hit_data <- hit_data %>% inner_join(grouped_pitching_stats, by = "pitcher_name")
```


Now prep for classification model with outcome column addition.
```{r}
hit_data <- hit_data %>% mutate(on_1b = ifelse(!is.na(on_1b), 1, 0), on_2b = ifelse(!is.na(on_2b), 1, 0), on_3b = ifelse(!is.na(on_3b), 1, 0))


hit_data <- hit_data %>% mutate(outcome = ifelse(description == "swinging_strike" | description == "swinging_strike_blocked" | description == "foul_tip" | (strikes == 2 & description == "foul_bunt") | description == "bunt_foul_tip" | description == "missed_bunt", "Miss", ifelse(description == "foul" | description == "foul_bunt", "Foul", ifelse(events == "single", "Single", ifelse(events == "double", "Double", ifelse(events == "triple", "Triple", ifelse(events == "home_run", "Homerun", "Out")))))))



```


Build classification model using xgboost to determine outcome probabilities of Miss, Foul, Out, Single, Double, Triple, and Homerun.
```{r}

swings <- hit_data %>% filter(description != "ball" & description != "called_strike" & description != "blocked_ball" & description != "pitchout" & description != "hit_by_pitch")

swings <- swings %>% filter(ERA <= 20)

set.seed(7777777)
sample <- sample(1:length(swings$game_date), length(swings$game_date) * .8)
train <- swings[sample,]
test <- swings[-sample,]

train_lab <- as.factor(train$outcome)
test_lab <- as.factor(test$outcome)

final_lab <- levels(train_lab)

train <- train %>% select(plate_x, plate_z, pfx_x, pfx_z, pitch_type, balls, strikes, outs_when_up, release_spin_rate, stand, p_throws, effective_speed, release_speed, PA, BA, OBP, SLG, barrel_perc, avg_exit_velocity, avg_launch_angle, xAVG, xwOBA, wOBA, ISO, IP, ERA, GB.FB, WHIP, SO_perc, uBB_perc, overall_spin_rate, overall_velocity, outcome)



test <- test %>% select(plate_x, plate_z, pfx_x, pfx_z, pitch_type, balls, strikes, outs_when_up, release_spin_rate, stand, p_throws, effective_speed, release_speed, PA, BA, OBP, SLG, barrel_perc, avg_exit_velocity, avg_launch_angle, xAVG, xwOBA, wOBA, ISO, IP, ERA, GB.FB, WHIP, SO_perc, uBB_perc, overall_spin_rate, overall_velocity, outcome)


train$pitch_type <- as.numeric(as.factor(train$pitch_type)) - 1
train$p_throws <- as.numeric(as.factor(train$p_throws)) - 1
train$stand <- as.numeric(as.factor(train$stand)) - 1


test$pitch_type <- as.numeric(as.factor(test$pitch_type)) - 1
test$p_throws <- as.numeric(as.factor(test$p_throws)) - 1
test$stand <- as.numeric(as.factor(test$stand)) - 1



train_lab <- as.numeric(as.factor(train_lab))-1
test_lab <- as.numeric(as.factor(test_lab))-1


xgb.train = xgb.DMatrix(data = as.matrix(train[, -33]), label = train_lab)
xgb.test = xgb.DMatrix(data = as.matrix(test[, -33]), label = test_lab)


##### Model
params <- list(objective = "multi:softprob", num_class = 7, eta = .3, min_child_weight = 6.47, max_depth = 9, subsample = .837, colsample_bytree = .75, eval_metric = "merror")


swing_mod <- xgb.train(params = params, data = xgb.train, nrounds = 49)
swing_mod
```


Remove unnecessary variables
```{r}
remove(ball_next, basic_batter_stats, basic_pitcher_stats, grouped_pitching_stats, grouped_stats, pitchers, pitchers_2019, pitchers_2020, run_expectancy_state_table, strike_next, takes_ball)

remove(train, test, takes, take_outcomes, swings)
```


Don't run this chunk, this is for the optimization of the parameters on the xgboost model which has already been done
```{r, eval = FALSE}
# Use built in cv to determine best value for nrounds
xgbcv <- xgb.cv(params = params, data = xgb.train, nrounds = 100, nfold = 5, showsd = T, stratified = T, print.every.n = 10, early.stop.round = 20, maximize = F)
##best iteration = 49

min(xgbcv$evaluation_log$train_merror_mean)
# Min error is  0.3959454

xgb.importance(feature_names = colnames(train),model = swing_mod)


#convert characters to factors
fact_col <- colnames(train)[sapply(train, is.character)]


train <- train[which(!is.na(train$barrel_perc)),]
train <- train[which(!is.na(train$xAVG)),]
train <- train[which(!is.na(train$ERA) & is.finite(train$ERA)),]
train <- train[which(!is.na(train$overall_spin_rate)),]

test <- test[which(!is.na(test$barrel_perc)),]
test <- test[which(!is.na(test$xAVG)),]
test <- test[which(!is.na(test$ERA) & is.finite(test$ERA)),]
test <- test[which(!is.na(test$overall_spin_rate)),]

#create tasks
traintask <- makeClassifTask(data = train, target = "outcome")
testtask <- makeClassifTask(data = test, target = "outcome")

#do one hot encoding`<br/> 
traintask <- createDummyFeatures (obj = traintask) 
testtask <- createDummyFeatures (obj = testtask)

#create learner
lrn <- makeLearner("classif.xgboost",predict.type = "response")
lrn$par.vals <- list( objective="multi:softprob", eval_metric="merror", nrounds=100L, eta=0.1)

#set parameter space
params <- makeParamSet( makeDiscreteParam("booster",values = c("gbtree","gblinear")), makeIntegerParam("max_depth",lower = 3L,upper = 10L), makeNumericParam("min_child_weight",lower = 1L,upper = 10L), makeNumericParam("subsample",lower = 0.5,upper = 1), makeNumericParam("colsample_bytree",lower = 0.5,upper = 1))

#set resampling strategy
rdesc <- makeResampleDesc("CV",stratify = T,iters=5L)

#search strategy
ctrl <- makeTuneControlRandom(maxit = 10L)

#set parallel backend
library(parallel)
library(parallelMap) 
parallelStartSocket(cpus = detectCores())

#parameter tuning
mytune <- tuneParams(learner = lrn, task = traintask, resampling = rdesc, measures = acc, par.set = params, control = ctrl, show.info = T)
mytune
mytune$y 
#0.4584903

#set hyperparameters
lrn_tune <- setHyperPars(lrn,par.vals = mytune$x)

#train model
xgmodel <- train(learner = lrn_tune,task = traintask)


#predict model
xgpred <- predict(xgmodel,testtask, type="response")

confusionMatrix(xgpred$data$response,xgpred$data$truth)
#Accuracy : 0.4609

########
xgb.pred = predict(swing_mod, xgb.test, reshape=T)
xgb.pred = as.data.frame(xgb.pred)
xgb.pred$label <- test$outcome
names(xgb.pred) <- levels(test$outcome)


# Use the predicted label with the highest probability
#xgb.pred$prediction = apply(xgb.pred,1,function(x) colnames(xgb.pred)[which.max(x)])
xgb.pred$label = swings$outcome[-sample]
colnames(xgb.pred)[1:7] <- final_lab

table(xgb.pred$label)



xgb.pred %>% group_by(label) %>% summarize(meanMiss = mean(Miss), 
                                           meanOut = mean(Out), 
                                           meanSingle = mean(Single), 
                                           meanDouble = mean(Double),
                                           meanTriple = mean(Triple),
                                           meanHR = mean(Homerun), 
                                           meanFoul = mean(Foul))
quantile(xgb.pred$Homerun, 0.99)
xgb.pred %>% filter(xgb.pred$Homerun > 0.075) %>% select(label) %>% table() %>% prop.table()
xgb.pred %>% select(label)%>% table() %>% prop.table()

#Check out results

xgb.pred %>% arrange(-Double)

```


## Test EAGLE on all pitches from years 2016-2019
Import data for each year and merge pitching data with hitting data.
```{r}
pitches_2016 <- read_rds("MLB Pitches 2016.rds")
pitches_2017 <- read_rds("MLB Pitches 2017.rds")
pitches_2018 <- read_rds("MLB Pitches 2018.rds")
pitches_2019 <- read_rds("MLB Pitches 2019.rds")

pitchers_2016 <- readRDS("Statcast Pitchers Pitch Data 2016.rds")
pitchers_2017 <- readRDS("Statcast Pitchers Pitch Data 2017.rds")
pitchers_2018 <- readRDS("Statcast Pitchers Pitch Data 2018.rds")
pitchers_2019 <- readRDS("Statcast Pitchers Pitch Data 2019.rds")

pitchers_2016 <- pitchers_2016 %>% select(game_pk, at_bat_number, pitch_number, player_name) %>% rename(pitcher_name = player_name)
pitches_2016 <- pitches_2016 %>% left_join(pitchers_2016, by = c("game_pk", "at_bat_number", "pitch_number"))

pitchers_2017 <- pitchers_2017 %>% select(game_pk, at_bat_number, pitch_number, player_name) %>% rename(pitcher_name = player_name)
pitches_2017 <- pitches_2017 %>% left_join(pitchers_2017, by = c("game_pk", "at_bat_number", "pitch_number"))

pitchers_2018 <- pitchers_2018 %>% select(game_pk, at_bat_number, pitch_number, player_name) %>% rename(pitcher_name = player_name)
pitches_2018 <- pitches_2018 %>% left_join(pitchers_2018, by = c("game_pk", "at_bat_number", "pitch_number"))

pitchers_2019 <- pitchers_2019 %>% select(game_pk, at_bat_number, pitch_number, player_name) %>% rename(pitcher_name = player_name)
pitches_2019 <- pitches_2019 %>% left_join(pitchers_2019, by = c("game_pk", "at_bat_number", "pitch_number"))


remove(pitchers_2016, pitchers_2017, pitchers_2018, pitchers_2019)

```


Make function to apply predictions from models and create EAGLE
```{r}
apply_EAGLE <- function(pitch_df, startdate, enddate){
  
  hit_data <- pitch_df
  startdate <- as.Date(startdate)
  enddate <- as.Date(enddate)
  
hit_data$pitch_id <- seq(1, nrow(hit_data))

hit_data <- hit_data %>% run_expectancy_code(level = "pitch")
hit_data$description <- as.factor(hit_data$description)



ball_next <- hit_data
ball_next <- ball_next %>% mutate(balls_new = ifelse(balls < 3, balls + 1, 0), 
                                  strikes_new = ifelse(balls < 3, strikes, 0), 
                                  outs_when_up_new = outs_when_up, 
                                  on_1b_new = ifelse(balls == 3 & is.na(on_1b), 1, on_1b), 
                                  on_2b_new = ifelse(balls == 3 & !is.na(on_1b), 1, on_2b), 
                                  on_3b_new = ifelse(balls == 3 & !is.na(on_1b) & !is.na(on_2b), 1, on_3b)) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
ball_next <- ball_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))

strike_next <- hit_data
strike_next <- strike_next %>% mutate(balls_new = ifelse(strikes < 2, balls, 0), 
                                  strikes_new = ifelse(strikes < 2, strikes + 1, 0), 
                                  outs_when_up_new = ifelse(strikes < 2, outs_when_up, outs_when_up + 1), 
                                  on_1b_new =  on_1b, 
                                  on_2b_new = on_2b, 
                                  on_3b_new = on_3b) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
strike_next <- strike_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))

strike_next[which(strike_next$outs_when_up_new == 3), "avg_re.y"] <- 0



#####
hit_data_ball <- ball_next %>% mutate(avg_re.ball = avg_re.y - avg_re.x) %>%
  select(pitch_id, balls_new, strikes_new, outs_when_up_new, on_1b_new, on_2b_new, on_3b_new, count_base_out_state_new, avg_re.ball) %>% 
  rename(count_base_out_state_ball = count_base_out_state_new)

hit_data_outcomes <- strike_next %>% mutate(avg_re.strike = avg_re.y - avg_re.x) %>%
  rename(count_base_out_state_orig = count_base_out_state, count_base_out_state_strike = count_base_out_state_new, avg_re.orig = avg_re.x) %>% left_join(hit_data_ball, by = "pitch_id")



hit_data_outcomes <- hit_data_outcomes %>% mutate(plate_x2 = plate_x^2, plate_z2 = plate_z^2,
         plate_x3 = plate_x2 * plate_x, plate_z3 = plate_z2 * plate_z,
         plate_xz = plate_x * plate_z,
         plate_x2z = plate_x^2 * plate_z,
         plate_xz2 = plate_x * plate_z^2)

hit_data_outcomes <- hit_data_outcomes %>% add_predictions(strike_prob_model, var = "strike_prob", type = "response")

hit_data_outcomes <- hit_data_outcomes %>% mutate(re_take = avg_re.ball * (1 - strike_prob) + avg_re.strike * strike_prob)


########
hit_data$pitch_id <- seq(1, nrow(hit_data))



###### Misses
miss_next <- hit_data
miss_next <- miss_next %>% mutate(balls_new = ifelse(strikes < 2, balls, 0), 
                                  strikes_new = ifelse(strikes < 2, strikes + 1, 0), 
                                  outs_when_up_new = ifelse(strikes < 2, outs_when_up, outs_when_up + 1), 
                                  on_1b_new = on_1b, 
                                  on_2b_new = on_2b, 
                                  on_3b_new = on_3b) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
miss_next <- miss_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))

miss_next[which(miss_next$outs_when_up_new == 3), "avg_re.y"] <- 0



###### Fouls

foul_next <- hit_data
foul_next <- foul_next %>% mutate(balls_new = balls, 
                                  strikes_new = ifelse(strikes < 2, strikes + 1, 2), 
                                  outs_when_up_new = outs_when_up, 
                                  on_1b_new = on_1b, 
                                  on_2b_new = on_2b, 
                                  on_3b_new = on_3b) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
foul_next <- foul_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))


###### Outs

out_next <- hit_data
out_next <- out_next %>% mutate(balls_new = 0, 
                                  strikes_new = 0, 
                                  outs_when_up_new = outs_when_up + 1, 
                                  on_1b_new = on_1b, 
                                  on_2b_new = on_2b, 
                                  on_3b_new = on_3b) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
out_next <- out_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))

out_next[which(out_next$outs_when_up_new == 3), "avg_re.y"] <- 0


###### Single

single_next <- hit_data
single_next <- single_next %>% mutate(balls_new = 0, 
                                  strikes_new = 0, 
                                  outs_when_up_new = outs_when_up, 
                                  on_1b_new = ifelse(is.na(on_1b), 1, on_1b), 
                                  on_2b_new = ifelse(is.na(on_1b), on_1b, 1), 
                                  on_3b_new = ifelse(is.na(on_2b), on_2b, 1)) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
single_next <- single_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))


####### Double

double_next <- hit_data
double_next <- double_next %>% mutate(balls_new = 0, 
                                  strikes_new = 0, 
                                  outs_when_up_new = outs_when_up, 
                                  on_1b_new = NA, 
                                  on_2b_new = 1, 
                                  on_3b_new = ifelse(is.na(on_1b), on_1b, 1)) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
double_next <- double_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))


###### Triple
triple_next <- hit_data
triple_next <- triple_next %>% mutate(balls_new = 0, 
                                  strikes_new = 0, 
                                  outs_when_up_new = outs_when_up, 
                                  on_1b_new = NA, 
                                  on_2b_new = NA, 
                                  on_3b_new = 1) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
triple_next <- triple_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))


####### Homerun

homerun_next <- hit_data
homerun_next <- homerun_next %>% mutate(balls_new = 0, 
                                  strikes_new = 0, 
                                  outs_when_up_new = outs_when_up, 
                                  on_1b_new = NA, 
                                  on_2b_new = NA, 
                                  on_3b_new = NA) %>% 
  mutate(count_base_out_state_new = paste(balls_new, "-", strikes_new, ", ", outs_when_up_new, " outs, ", ifelse(!is.na(.$on_1b_new), "1b", "_"), ifelse(!is.na(.$on_2b_new), "2b", "_"), ifelse(!is.na(.$on_3b_new), "3b", "_")))
# merge run exxpectancy table to new data
homerun_next <- homerun_next %>% left_join(run_expectancy_state_table, by = c("count_base_out_state_new" = "count_base_out_state"))

########

hit_data_miss <- miss_next %>% mutate(avg_re.miss = avg_re.y - avg_re.x) %>%
  select(pitch_id, balls_new, strikes_new, outs_when_up_new, on_1b_new, on_2b_new, on_3b_new, count_base_out_state_new, avg_re.miss) %>% 
  rename(count_base_out_state_miss = count_base_out_state_new)

hit_data_foul <- foul_next %>% mutate(avg_re.foul = avg_re.y - avg_re.x) %>%
  select(pitch_id, balls_new, strikes_new, outs_when_up_new, on_1b_new, on_2b_new, on_3b_new, count_base_out_state_new, avg_re.foul) %>% 
  rename(count_base_out_state_foul = count_base_out_state_new)

hit_data_out <- out_next %>% mutate(avg_re.out = avg_re.y - avg_re.x) %>%
  select(pitch_id, balls_new, strikes_new, outs_when_up_new, on_1b_new, on_2b_new, on_3b_new, count_base_out_state_new, avg_re.out) %>% 
  rename(count_base_out_state_out = count_base_out_state_new)

hit_data_single <- single_next %>% mutate(avg_re.single = avg_re.y - avg_re.x) %>%
  select(pitch_id, balls_new, strikes_new, outs_when_up_new, on_1b_new, on_2b_new, on_3b_new, count_base_out_state_new, avg_re.single) %>% 
  rename(count_base_out_state_single = count_base_out_state_new)

hit_data_double <- double_next %>% mutate(avg_re.double = avg_re.y - avg_re.x) %>%
  select(pitch_id, balls_new, strikes_new, outs_when_up_new, on_1b_new, on_2b_new, on_3b_new, count_base_out_state_new, avg_re.double) %>% 
  rename(count_base_out_state_double = count_base_out_state_new)

hit_data_triple <- triple_next %>% mutate(avg_re.triple = avg_re.y - avg_re.x) %>%
  select(pitch_id, balls_new, strikes_new, outs_when_up_new, on_1b_new, on_2b_new, on_3b_new, count_base_out_state_new, avg_re.triple) %>% 
  rename(count_base_out_state_triple = count_base_out_state_new)

hit_data_homerun <- homerun_next %>% mutate(avg_re.homerun = avg_re.y - avg_re.x) %>%
  select(pitch_id, balls_new, strikes_new, outs_when_up_new, on_1b_new, on_2b_new, on_3b_new, count_base_out_state_new, avg_re.homerun) %>% 
  rename(count_base_out_state_homerun = count_base_out_state_new)



hit_data_outcomes <- hit_data_outcomes %>%
  left_join(hit_data_miss, by = "pitch_id") %>% 
  left_join(hit_data_foul, by = "pitch_id") %>% 
  left_join(hit_data_out, by = "pitch_id") %>% 
  left_join(hit_data_single, by = "pitch_id") %>% 
  left_join(hit_data_double, by = "pitch_id") %>% 
  left_join(hit_data_triple, by = "pitch_id") %>% 
  left_join(hit_data_homerun, by = "pitch_id")

hit_data_outcomes <- hit_data_outcomes %>% 
  mutate(avg_re.ball = ifelse(!is.na(on_1b) & !is.na(on_2b) & !is.na(on_3b) & balls == 3, avg_re.ball + 1, avg_re.ball)) %>% 
  mutate(avg_re.single = ifelse(is.na(on_3b), avg_re.single, avg_re.single + 1)) %>% 
  mutate(avg_re.double = ifelse(!is.na(on_3b) & !is.na(on_2b), avg_re.double + 2, ifelse(!is.na(on_3b) | !is.na(on_2b), avg_re.double + 1, avg_re.double))) %>% 
  mutate(avg_re.triple = ifelse(!is.na(on_3b) & !is.na(on_2b) & !is.na(on_1b), avg_re.triple + 3, ifelse((!is.na(on_3b) & !is.na(on_2b)) | (!is.na(on_1b) & !is.na(on_2b)) | (!is.na(on_3b) & !is.na(on_1b)), avg_re.triple + 2, ifelse(!is.na(on_1b) | !is.na(on_2b) | !is.na(on_3b), avg_re.triple + 1, avg_re.triple)))) %>%
  mutate(avg_re.homerun = ifelse(!is.na(on_3b) & !is.na(on_2b) & !is.na(on_1b), avg_re.triple + 4, ifelse((!is.na(on_3b) & !is.na(on_2b)) | (!is.na(on_1b) & !is.na(on_2b)) | (!is.na(on_3b) & !is.na(on_1b)), avg_re.triple + 3, ifelse(!is.na(on_1b) | !is.na(on_2b) | !is.na(on_3b), avg_re.homerun + 2, 1 + avg_re.homerun))))



########
## merge to data 

basic_batter_stats <- daily_batter_bref(startdate, enddate)

basic_batter_stats$bbref_id <- as.numeric(basic_batter_stats$bbref_id)

hit_data_outcomes <- hit_data_outcomes %>% separate(player_name, into = c("Last", "First"), sep = ", ") %>% mutate(Name = paste(First, Last))


hit_data_outcomes <- hit_data_outcomes %>% inner_join(basic_batter_stats, by = c("Name"))



hit_data_outcomes <- hit_data_outcomes %>% mutate(barrel = ifelse(launch_angle <= 50 & launch_speed >= 98 & launch_speed * 1.5 - launch_angle >= 117 & launch_speed + launch_angle >= 124, 1, 0))


# Fix character columns
hit_data_outcomes[which(hit_data_outcomes$launch_speed_angle == "null"),"launch_speed_angle"] <- NA
hit_data_outcomes[which(hit_data_outcomes$estimated_ba_using_speedangle == "null"), "estimated_ba_using_speedangle"] <- NA
hit_data_outcomes[which(hit_data_outcomes$estimated_woba_using_speedangle == "null"), "estimated_woba_using_speedangle"] <- NA
hit_data_outcomes[which(hit_data_outcomes$woba_value == "null"), "woba_value"] <- NA
hit_data_outcomes[which(hit_data_outcomes$iso_value == "null"), "iso_value"] <- NA

hit_data_outcomes$launch_speed_angle <- as.numeric(hit_data_outcomes$launch_speed_angle)
hit_data_outcomes$estimated_ba_using_speedangle <- as.numeric(hit_data_outcomes$estimated_ba_using_speedangle)
hit_data_outcomes$estimated_woba_using_speedangle <- as.numeric(hit_data_outcomes$estimated_woba_using_speedangle)
hit_data_outcomes$woba_value <- as.numeric(hit_data_outcomes$woba_value)
hit_data_outcomes$iso_value <- as.numeric(hit_data_outcomes$iso_value)


grouped_stats <- hit_data_outcomes %>% group_by(batter, zone) %>% summarise(barrel_perc = mean(barrel, na.rm = T), avg_launch_angle = mean(launch_angle, na.rm = T), avg_exit_velocity = mean(launch_speed, na.rm = T), avg_launch_speed_angle = mean(launch_speed_angle, na.rm = T), xAVG = mean(estimated_ba_using_speedangle, na.rm = T), xwOBA = mean(estimated_woba_using_speedangle, na.rm = T), wOBA = mean(woba_value, na.rm = T), ISO = mean(iso_value, na.rm = T))



hit_data_outcomes <- hit_data_outcomes %>% inner_join(grouped_stats, by = c("batter", "zone"))

hit_data_outcomes <- hit_data_outcomes %>% separate(pitcher_name, into = c("last_name", "first_name"), sep = ", ") %>% mutate(pitcher_name = paste(first_name, last_name, sep = " "))

basic_pitcher_stats <- daily_pitcher_bref(startdate, enddate)

basic_pitcher_stats$bbref_id <- as.numeric(basic_pitcher_stats$bbref_id)

hit_data_outcomes <- hit_data_outcomes %>% inner_join(basic_pitcher_stats, by = c("pitcher_name" = "Name"))

grouped_pitching_stats <- hit_data_outcomes %>% group_by(pitcher) %>% summarise(FF_spin_rate = mean(release_spin_rate[pitch_type == "FF"], na.rm = T), CU_spin_rate = mean(release_spin_rate[pitch_type == "CU"], na.rm = T), FF_velocity = mean(release_speed[pitch_type == "FF"], na.rm = T), BRK_velocity = mean(release_speed[pitch_type == "CU" | pitch_type == "SL"], na.rm = T), overall_spin_rate = mean(release_spin_rate, na.rm = T), overall_velocity = mean(release_speed, na.rm = T))


hit_data_outcomes <- hit_data_outcomes %>% inner_join(grouped_pitching_stats, by = "pitcher")


########
hit_data_outcomes <- hit_data_outcomes %>% mutate(outcome = ifelse(description == "swinging_strike" | description == "swinging_strike_blocked" | description == "foul_tip" | (strikes == 2 & description == "foul_bunt") | description == "bunt_foul_tip" | description == "missed_bunt", "Miss", ifelse(description == "foul" | description == "foul_bunt", "Foul", ifelse(events == "single", "Single", ifelse(events == "double", "Double", ifelse(events == "triple", "Triple", ifelse(events == "home_run", "Homerun", "Out")))))))

hit_data_outcomes <- hit_data_outcomes %>% filter(ERA <= 20)


all_data <- hit_data_outcomes %>% select(plate_x, plate_z, pfx_x, pfx_z, pitch_type, balls, strikes, outs_when_up, release_spin_rate, stand, p_throws, effective_speed, release_speed, PA, BA, OBP, SLG, barrel_perc, avg_exit_velocity, avg_launch_angle, xAVG, xwOBA, wOBA, ISO, IP, ERA, GB.FB, WHIP, SO_perc, uBB_perc, overall_spin_rate, overall_velocity, outcome)


all_lab <- as.factor(all_data$outcome)
all_lab <- as.numeric(as.factor(all_lab))-1

all_data$pitch_type <- as.numeric(as.factor(all_data$pitch_type)) - 1
all_data$p_throws <- as.numeric(as.factor(all_data$p_throws)) - 1
all_data$stand <- as.numeric(as.factor(all_data$stand)) - 1


xgb.all = xgb.DMatrix(data = as.matrix(all_data[, -33]), label = all_lab)

## Predict
all.pred = predict(swing_mod, xgb.all, reshape=T)
all.pred = as.data.frame(all.pred)
names(all.pred) <- levels(all_data$outcome)
all.pred$label <- all_data$outcome


# Use the predicted label with the highest probability
#all.pred$prediction = apply(all.pred,1,function(x) colnames(all.pred)[which.max(x)])
colnames(all.pred)[1:7] <- final_lab


hit_data_outcomes <- cbind(hit_data_outcomes, all.pred)





# Calculate RE of a swing

hit_data_outcomes <- hit_data_outcomes %>% mutate(re_swing = avg_re.miss * Miss + avg_re.foul * Foul + avg_re.out * Out + avg_re.single * Single + avg_re.double * Double + avg_re.triple * Triple + avg_re.homerun * Homerun)



#Now create the stat to evaluate players on
#Expected Runs Added by swinging

hit_data_outcomes <- hit_data_outcomes %>% mutate(expRA_swing = re_swing - re_take, swing = ifelse(description == "ball" | description == "called_strike" | description == "blocked_ball" | description == "pitchout", 0, 1))

hit_data_outcomes <- hit_data_outcomes %>% mutate(abs_eRA_s = abs(expRA_swing), should_swing = ifelse(re_swing > re_take, 1, 0), correct_choice = ifelse(should_swing == swing, 1, 0))

hit_data_outcomes <- hit_data_outcomes %>% mutate(runs_lost_bad_dec = ifelse(correct_choice == 0, abs_eRA_s, 0))

hit_data_outcomes
}


```



Apply the EAGLE function to the data
```{r, warning=F}
remove(hit_data)
data_2016 <- apply_EAGLE(pitches_2016, startdate = "2016-04-03", enddate = "2016-10-02")
remove(pitches_2016)
data_2017 <- apply_EAGLE(pitches_2017, startdate = "2017-04-02", enddate = "2017-10-01")
remove(pitches_2017)
data_2018 <- apply_EAGLE(pitches_2018, startdate = "2018-03-29", enddate = "2018-09-30")
remove(pitches_2018)
data_2019 <- apply_EAGLE(pitches_2019, startdate = "2019-03-20", enddate = "2019-09-29")
remove(pitches_2019)

```




Create combined season data frame with commonly used metrics and EAGLE so that we have each hitter from 2016-2019
```{r}
eye_2016 <- data_2016 %>% filter(PA >= 200, !is.na(expRA_swing)) %>% group_by(batter, Name) %>% summarize(year = first(season.x), runs_lost = sum(runs_lost_bad_dec) / sum(abs_eRA_s), wRA = sum(ifelse(swing == 1, expRA_swing, -expRA_swing)), EAGLE = wRA / n(), avg = first(BA), obp = first(OBP), ops = first(OPS),  kperc = first(SO.x) / first(PA), bbperc = first(BB.x) / first(PA), perc_swing_out_zone = sum(ifelse(strike_prob < .3 & swing == 1, 1, 0)) / sum(ifelse(strike_prob < .3, 1, 0))) 

eye_2017 <- data_2017 %>% filter(PA >= 200, !is.na(expRA_swing)) %>% group_by(batter, Name) %>% summarize(year = first(season.x), runs_lost = sum(runs_lost_bad_dec) / sum(abs_eRA_s), wRA = sum(ifelse(swing == 1, expRA_swing, -expRA_swing)), EAGLE = wRA / n(), avg = first(BA), obp = first(OBP), ops = first(OPS),  kperc = first(SO.x) / first(PA), bbperc = first(BB.x) / first(PA), perc_swing_out_zone = sum(ifelse(strike_prob < .3 & swing == 1, 1, 0)) / sum(ifelse(strike_prob < .3, 1, 0))) 

eye_2018 <- data_2018 %>% filter(PA >= 200, !is.na(expRA_swing)) %>% group_by(batter, Name) %>% summarize(year = first(season.x), runs_lost = sum(runs_lost_bad_dec) / sum(abs_eRA_s), wRA = sum(ifelse(swing == 1, expRA_swing, -expRA_swing)), EAGLE = wRA / n(), avg = first(BA), obp = first(OBP), ops = first(OPS), kperc = first(SO.x) / first(PA), bbperc = first(BB.x) / first(PA), perc_swing_out_zone = sum(ifelse(strike_prob < .3 & swing == 1, 1, 0)) / sum(ifelse(strike_prob < .3, 1, 0))) 

eye_2019 <- data_2019 %>% filter(PA >= 200, !is.na(expRA_swing)) %>% group_by(batter, Name) %>% summarize(year = first(season.x), runs_lost = sum(runs_lost_bad_dec) / sum(abs_eRA_s), wRA = sum(ifelse(swing == 1, expRA_swing, -expRA_swing)), EAGLE = wRA / n(), avg = first(BA), obp = first(OBP), ops = first(OPS), kperc = first(SO.x) / first(PA), bbperc = first(BB.x) / first(PA), perc_swing_out_zone = sum(ifelse(strike_prob < .3 & swing == 1, 1, 0)) / sum(ifelse(strike_prob < .3, 1, 0))) 


eye_all <- eye_2016 %>% full_join(eye_2017, by = c("batter")) %>% full_join(eye_2018, by = c("batter")) %>% full_join(eye_2019, by = c("batter"))

eye_all <- eye_all %>% rename(wRL_2016 = runs_lost.x, wRL_2017 = runs_lost.y, wRL_2018 = runs_lost.x.x, wRL_2019 = runs_lost.y.y)

```


Create data frame containing player stats from consecutive years so that we can see predictabililty and correlation across years
```{r}
eye_this_next1 <- eye_2016 %>% left_join(eye_2017, by = c("batter", "Name")) 
eye_this_next2 <- eye_2017 %>% left_join(eye_2018, by = c("batter", "Name"))
eye_this_next3 <- eye_2018 %>% left_join(eye_2019, by = c("batter", "Name"))

eye_this_next <- bind_rows(eye_this_next1, eye_this_next2, eye_this_next3)

remove(eye_this_next1, eye_this_next2, eye_this_next3)

```


Look at results of outcome probabilities xgboost model
```{r}
all_years <- rbind(data_2016, data_2017, data_2018, data_2019)

levs = c("Miss", "Foul", "Out", "Single", "Double", "Triple", "Homerun")

all_years %>% 
  filter(!is.na(outcome)) %>% 
  group_by(outcome) %>% 
  summarize(Miss = mean(Miss, na.rm = T), 
            Foul = mean(Foul, na.rm = T), 
            Out = mean(Out, na.rm = T), 
            Single = mean(Single, na.rm = T), 
            Double = mean(Double, na.rm = T), 
            Triple = mean(Triple, na.rm = T), 
            Homerun = mean(Homerun, na.rm = T)) %>% 
  mutate(outcome =  factor(outcome, levels = levs)) %>%
  arrange(outcome)

```


Run linear regression models in order to test for significant relationships between EAGLE and other significant statistics
```{r}

####
# Test EAGLE correlation from year to year
eye_this_next_mod <- lm(EAGLE.y ~ EAGLE.x, data = eye_this_next)
summary(eye_this_next_mod)

####

# Test with AVG
ba_eye_mod <- lm(avg.x ~ EAGLE.x, data = eye_this_next)
summary(ba_eye_mod)

next_ba_eye_mod <- lm(avg.y ~ EAGLE.x, eye_this_next)
summary(next_ba_eye_mod)

both_ba_eye_mod <- lm(avg.y ~ EAGLE.x + avg.x, eye_this_next)
summary(both_ba_eye_mod)

# Test for OBP
obp_eye_mod <- lm(obp.x ~ EAGLE.x, data = eye_this_next)
summary(obp_eye_mod)

next_obp_eye_mod <- lm(obp.y ~ EAGLE.x, eye_this_next)
summary(next_obp_eye_mod)

both_obp_eye_mod <- lm(obp.y ~ EAGLE.x + obp.x, eye_this_next)
summary(both_obp_eye_mod)

# Now for OPS
ops_eye_mod <- lm(ops.x ~ EAGLE.x, data = eye_this_next)
summary(ops_eye_mod)

next_ops_eye_mod <- lm(ops.y ~ EAGLE.x, eye_this_next)
summary(next_ops_eye_mod)

both_ops_eye_mod <- lm(ops.y ~ EAGLE.x + ops.x, eye_this_next)
summary(both_ops_eye_mod)

###
```


#Create graphs for important relationships

First create graph for walk percentage while also testing the relationship with a linear regression model
```{r}
eye_this_next %>% filter(EAGLE.x >= 0) %>% ggplot(aes(x = EAGLE.x, y = bbperc.x)) + 
  geom_point() +
  stat_smooth(method = "lm") + 
  theme_gdocs() + 
  labs(title = "Relationship between EAGLE and BB%", 
       subtitle = "2016-2019") +
  xlab("EAGLE") +
  ylab("BB%")

bbperc_mod <- lm(bbperc.x ~ EAGLE.x, data = eye_this_next)
summary(bbperc_mod)

```

Graph for relationship with O% and also test it with a linear regression model
```{r}
eye_this_next %>% filter(EAGLE.x >= 0) %>% ggplot(aes(x = EAGLE.x, y = perc_swing_out_zone.x)) + 
  geom_point() +
  stat_smooth(method = "lm") + 
  theme_gdocs() + 
  labs(title = "Relationship between EAGLE and O%", 
       subtitle = "2016-2019") +
  xlab("EAGLE") +
  ylab("O%")

operc_mod <- lm(perc_swing_out_zone.x ~ EAGLE.x, data = eye_this_next)
summary(operc_mod)
```


Create graph for relationship of EAGLE with OBP
```{r}
eye_this_next %>% 
  filter(!is.na(EAGLE.x), !is.na(obp.x), EAGLE.x >= .025) %>% 
  ggplot(aes(EAGLE.x, obp.x)) + 
  geom_point() + 
  stat_smooth(method = "lm") + 
  theme_gdocs() + 
  ggtitle("Correlation to OBP") + 
  xlab("EAGLE") + 
  ylab("OBP")

```


Create graph to highlight the correlation of EAGLE from year to year
```{r}
eye_this_next %>% filter(EAGLE.x >= 0) %>% ggplot(aes(x = EAGLE.x, y = EAGLE.y)) + 
  geom_point() +
  stat_smooth(method = "lm") + 
  theme_gdocs() + 
  labs(title = "Relationship between EAGLEs from year to year", 
       subtitle = "2016-2019") +
  xlab("EAGLE 1") +
  ylab("EAGLE 2")

```

Create graphs showing the relationship of EAGLE with the current year's OPS and with the next year's OPS
```{r}
eye_this_next %>% filter(EAGLE.x >= 0) %>% ggplot(aes(x = EAGLE.x, y = ops.x)) + 
  geom_point() +
  stat_smooth(method = "lm") + 
  theme_gdocs() + 
  labs(title = "Relationship between EAGLE and OPS", 
       subtitle = "2016-2019") +
  xlab("EAGLE") +
  ylab("OPS")

eye_this_next %>% filter(EAGLE.x >= 0) %>% ggplot(aes(x = EAGLE.x, y = ops.y)) + 
  geom_point() +
  stat_smooth(method = "lm") + 
  theme_gdocs() + 
  labs(title = "Relationship between EAGLE and Next Year's OPS", 
       subtitle = "2016-2019") +
  xlab("EAGLE") +
  ylab("Next Year's OPS")
```


Bring the data of all the years together to calculate predictions from percentiles in addition to testing correlation. Also check out who is the best/worst in EAGLE
```{r}

all_eye <- all_years %>% 
  mutate(season.x = as.numeric(season.x)) %>% 
  filter(PA >= 300, !is.na(expRA_swing)) %>% 
  group_by(Name, season.x) %>% 
  summarize(runs_lost = sum(runs_lost_bad_dec) / sum(abs_eRA_s), 
            wRA = sum(ifelse(swing == 1, expRA_swing, -expRA_swing)),
            ba = first(BA),
            obp = first(OBP),
            slg = first(SLG),
            ops = first(OPS),
            pitches = n(),
            bb_perc = first(BB.x) / first(PA),
            o_perc = sum(ifelse(strike_prob < .5 & swing == 1, 1, 0), na.rm = T) / sum(ifelse(strike_prob < .5, 1, 0), na.rm = T)) %>% 
  mutate(EAGLE = wRA / pitches)

all_eye %>% select(Name, season.x, EAGLE, bb_perc, o_perc) %>% arrange(-EAGLE) %>% head()
all_eye %>% select(Name, season.x, EAGLE, bb_perc, o_perc) %>% arrange(EAGLE) %>% head()

eagle_perc <- quantile(all_eye$EAGLE, probs = c(.1, .25, .75, .9))
percentiles <- data_frame(percentile = c("10th percentile", "25th percentile", "75th percentile", "90th percentile"), EAGLE.x = eagle_perc)
percentiles %>% add_predictions(ops_eye_mod, var = "OPS")
percentiles %>% add_predictions(next_ops_eye_mod, var = "Next_OPS")


cor(all_eye$EAGLE, all_eye$ops)

cor(all_eye$o_perc, all_eye$ops)
```

Current Year EAGLE correlated with current year statistics and next year's EAGLE
```{r}
eye_this_next %>% 
  ungroup() %>% 
  select(EAGLE.x, avg.x, obp.x, ops.x, kperc.x, bbperc.x, perc_swing_out_zone.x, EAGLE.y) %>% 
  rename(o_perc = perc_swing_out_zone.x, EAGLE_cur_year = EAGLE.x, EAGLE_next_year = EAGLE.y) %>% 
  cor(use = "complete.obs")

```

